HyperLogLog: Exponentially Bad in Adversarial Settings

### 2020-05-25 docs/HLL_exp_bad/main.tex
fix up
pseudocode

This simple approach suffers from large variance in the estimation. In order to improve the estimate, we can use many estimators in parallel instead of one, and average the results. Each estimate is stored in its own bucket. The collection of buckets is referred to as the HLL \emph{sketch}. To map a value $x$ to one of $2^n$ buckets, \cite{loglog} suggests using the first $n$ bits of $h(x)$ as a bucket index, and then computing the longest sequence of leading zero bits on the remaining $\ell$ bits of $h(x)$ (so that the output length of $h(\cdot)$ is $n+\ell$). Estimators from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha$, where $\alpha$ is empirically computed to correct a systematic multiplicative bias. As result, HLL is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using only 1.5 kB of memory~\cite{hll}.

@@

@@

\begin{table}[h]
\caption{Attack results, averaged over 30 iterations.}
\begin{tabular}{| m{8.5em} | m{4em} | m{4em} | m{4em} | m{4em} |}
    \hline
    \textbf{Scenario} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
    \textbf{Initial data?} & yes & no & yes & no \\ \hline
    \textbf{Original Card. Est.} & 1,000 & 0 & 1,005 & 0 \\ \hline
    \textbf{\# Items added} & 3,655,744 & 3,655,740 & 6,090,003 & 14,220 \\ \hline
    \textbf{Final Card. Est.} & 1'011 & 367 & 1,005 & 1 \\ \hline
\end{tabular}
\label{table:tab1}
\end{table}

\begin{table}[h]
\caption{Attack results, averaged over 30 iterations, in the setting of \cite{hllvuln}.}
\begin{tabular}{| m{8.5em} | m{4em} | m{4em} | m{4em} | m{4em} |}
    \hline
    \textbf{Scenario} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
    \textbf{Initial data?} & yes & no & yes & no \\ \hline
    \textbf{Original Card. Est.} & 1,004 & 0 & 998 & 0 \\ \hline
    \textbf{\# Items added} & 124,940 & 125,085 & 204,669 & 501 \\ \hline
    \textbf{Final Card. Est.} & 1,058 & 367 & 998 & 1 \\ \hline
\end{tabular}
\label{table:tab2}
\end{table}

@@

A formal analysis of using salting as a countermeasure to our attacks remains open.
%Future work could consist in finding mitigations that do not rely on salt. For example, would the imposed ID setting (like Facebook) solve the issue but still be convenient for the various HLL applications?
Also left for future work is the task of extending our attacks to HLL variants such as HLL++~\cite{hllpratice} and sliding HLL~\cite{slidinghll}.

>>>>

This simple approach suffers from large variance in the estimation. In order to improve the estimate, we can use many estimators in parallel instead of one, and average the results. Each estimate is stored in its own bucket. The collection of buckets is referred to as the HLL \emph{sketch}. To map a value $x$ to one of $2^n$ buckets, \cite{loglog} suggests using the first $n$ bits of $h(x)$ as a bucket index, and then computing the longest sequence of leading zero bits on the remaining $\ell$ bits of $h(x)$ (so that the output length of $h(\cdot)$ is $n+\ell$). Estimators from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha$, where $\alpha$ is empirically computed to correct a systematic multiplicative bias. On top of that, correction for small and big range are applied as shown in \ref{fig:hll}. As result, HLL is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using only 1.5 kB of memory~\cite{hll}.

@@

\begin{figure}[h]
\centering
\fbox{
\begin{minipage}{40em}
\begin{flushleft}
\textit{Let $h : D \rightarrow \{0, 1\}^{32}$ hash data from $\mathcal{D}$ to binary $32$–bit words.\\
Let $\rho(s)$ be the position of the leftmost $1$-bit of $s$: e.g., $\rho(1...) = 1, \rho(0001...) = 4, \rho(0^K) = K + 1.$\\}
\textbf{define} $\alpha_{16} = 0.673$; $\alpha_{32} = 0.697$; $\alpha_{64} = 0.709$; $\alpha_{m} = 0.7213/(1+ 1.079/m)$ for $m \ge 128$;\\
\vspace{0.5em}
\textbf{Program} HyperLogLog (\textbf{input} $\mathcal{M}$ : multiset of items from domain $\mathcal{D}$).\\
\textbf{assume} $m = 2^b$ with $b \in [4..16].$\\
\textbf{initialize} a collection of $m$ registers, $M[1], ..., M[m]$, to $0$;\\
\vspace{0.5em}
\textbf{for} $v \in \mathcal{M}$ \textbf{do}\\
    \hspace{2em} \textbf{set} $x := h(v);$\\
    \hspace{2em} \textbf{set} $j = 1 + <x_1x_2...x_{b}>_2$; \hfill \textit{\{the binary address determined by the first $b$ bits of $x$\}}\\
    \hspace{2em} \textbf{set} $w := x_{b+1}x_{b+2}...$;\\
   \hspace{2em} \textbf{set} $M[j] :=$ max$(M[j], \rho(w))$;\\
\vspace{0.2em}
\textbf{compute} $E := \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}$; \hfill \textit{\{the “raw” HyperLogLog estimate\}}\\
\vspace{0.2em}
\textbf{if} $E \le \frac{5}{2}m$ \textbf{then}\\
    \hspace{2em}  \textbf{let} $V$ be the number of registers equal to $0$;\\
    \hspace{2em}  \textbf{if} $V \neq 0$ \textbf{then set} $E^*:= m $log$(m/V)$ \textbf{else set} $E^*:= E$; \hfill \textit{\{small range correction\}}\\
\textbf{if} $E \le \frac{1}{30}2^{32}$ \textbf{then}\\
    \hspace{2em} \textbf{set} $E^*:= E$; \hfill \textit{\{intermediate range—no correction\}}\\
\textbf{if} $E > \frac{1}{30}2^{32}$ \textbf{then}\\
    \hspace{2em}  \textbf{set} $E^*:= -2^{32}$log$(1 - E/2^{32})$; \hfill \textit{\{large range correction\}}\\
\textbf{return} \textit{cardinality estimate $E^*$ with typical relative error} $\pm1.04/\sqrt{m}$.
\end{flushleft}
\end{minipage}
}
\caption{The HyperLogLog Program}
\label{fig:hll}
\end{figure}

@@

\begin{table}[h]
\centering
\caption{Attack results, averaged over 30 iterations.}
\begin{tabular}{| m{8.5em} | m{4em} | m{4em} | m{4em} | m{4em} |}
    \hline
    \textbf{Scenario} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
    \textbf{Initial data?} & yes & no & yes & no \\ \hline
    \textbf{Original Card. Est.} & 1,000 & 0 & 1,005 & 0 \\ \hline
    \textbf{\# Items added} & 3,655,744 & 3,655,740 & 6,090,003 & 14,220 \\ \hline
    \textbf{Final Card. Est.} & 1'011 & 367 & 1,005 & 1 \\ \hline
\end{tabular}
\label{table:tab1}
\end{table}

\begin{table}[h]
\centering
\caption{Attack results, averaged over 30 iterations, in the setting of \cite{hllvuln}.}
\begin{tabular}{| m{8.5em} | m{4em} | m{4em} | m{4em} | m{4em} |}
    \hline
    \textbf{Scenario} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
    \textbf{Initial data?} & yes & no & yes & no \\ \hline
    \textbf{Original Card. Est.} & 1,004 & 0 & 998 & 0 \\ \hline
    \textbf{\# Items added} & 124,940 & 125,085 & 204,669 & 501 \\ \hline
    \textbf{Final Card. Est.} & 1,058 & 367 & 998 & 1 \\ \hline
\end{tabular}
\label{table:tab2}
\end{table}

@@

A formal analysis of using salting as a countermeasure to our attacks remains open.
%Future work could consist in finding mitigations that do not rely on salt. For example, would the imposed ID setting (like Facebook) solve the issue but still be convenient for the various HLL applications?
Also left for future work is the task of extending our attacks to HLL variants such as redis alternative using maximum likelihood~\cite{newhll}, HLL++~\cite{hllpratice} and sliding HLL~\cite{slidinghll}.
